\documentclass[conference]{IEEEtran}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{url}
\usepackage{tikz}
\usetikzlibrary{positioning}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{axiom}{Axiom}

\begin{document}

\title{SVCA: Self-Verifying Computational Artifacts \\
\large A Primitive for Artifact-Oriented Science}

\author{\IEEEauthorblockN{Mateus}
\IEEEauthorblockA{MatVerse Lab\\
\{mateus\}@matverse.org}}

\maketitle

\begin{abstract}
Modern computational science lacks a \emph{canonical unit of trust}: every experiment relies on mutable infrastructure, external registries and human-managed links.
We formalize \emph{Self-Verifying Computational Artifacts} (SVCA) — digital objects that \emph{are} the experiment, cryptographically prove their own validity and reproduce byte-identically without infrastructure.
We axiomatise SVCA as the \emph{minimal class} closed under five invariants and prove that any object in this class yields \emph{self-sufficient evidence}.
\end{abstract}

\section{Introduction}
Computational science is built on a fragile premise: that \emph{descriptions} remain faithful to \emph{executions}.
We show this is not a social failure but an \emph{ontological mismatch}: papers are approximations, while reality is executable.
We introduce \textbf{Self‑Verifying Computational Artifacts} (SVCA) as the missing primitive.

\section{Axiomatic Definition}
An artifact \(A\) belongs to SVCA iff it satisfies:
\begin{axiom}[Integrity]
\(\exists h=\text{Hash}(A)\) (negligible collision) invariant under transport.
\end{axiom}
\begin{axiom}[Executability]
\(\exists E: E(A)\to S\) (valid state).
\end{axiom}
\begin{axiom}[Public Verifiability]
\(\exists V(pk,h,\sigma)\) with \(V(pk,h,\sigma)=1 \iff \sigma=\text{Sign}(sk,h)\).
\end{axiom}
\begin{axiom}[Strong Reproducibility]
\(\text{Build}(L) \equiv A\) byte-identical for all replicas.
\end{axiom}
\begin{axiom}[Autonomy]
\(A\) requires no external resources (DNS, registry, clock).
\end{axiom}
Removal of any axiom breaks the class — SVCA is minimal.

\begin{theorem}[Self‑Sufficient Evidence]
If \(A\in\text{SVCA}\), experimental validity is decidable offline.
\end{theorem}
\begin{proof}
By A3, \(\sigma\) exists with \(V(pk,\text{Hash}(A),\sigma)=1\). By A4, any observer recomputes \(A'\) and verifies \(\text{Hash}(A')=\text{Hash}(A)\). Decision reduces to pure mathematics.
\end{proof}
\begin{corollary}
Trust shifts from institutions to mathematics.
\end{corollary}

\section{Implementation and Evaluation}
We built a compact capsule containing a WASM module compiled from Go 1.22.3 with `-trimpath -ldflags="-buildid="`.
The build is containerized with a pinned Debian digest; toolchains are verified by SHA256.

\section{Implications}
\paragraph{Constitutional Computation.}
Execution occurs iff invariants hold — governance by cryptographic law.
\paragraph{Scientific Permanence.}
As long as a single copy exists, integrity can be re‑established without authorities.
\paragraph{Epistemic Shift.}
SVCA is to computational science what the specimen is to biology: the object itself is the evidence.

\section{Conclusion}
An experiment that cannot be independently re‑instantiated is not evidence — it is testimony.

\end{document}
